The brightness and spatial distributions of terrestrial radio sources

### 2013-04-01 articles/Article-RFI-distribution/article-rfi-distribution.tex
Work on ref report

\begin{abstract}
Faint undetected sources of radio-frequency interference (RFI) might become visible after long-time integration of radio observations if they are consistently present over time. Thereby, they might obstruct the detection of the weak astronomical signals of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals from the time of the earliest structures in the Universe. We explore the RFI situation at 30--163~MHz by studying brightness histograms of visibility data observed with LOFAR, similar to radio-source-count analyses that are used in cosmology. The histograms show an RFI distribution that follows a power-law distribution with an estimated exponent around -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law implies that the current LOFAR EoR observations should be severely RFI limited if the integrated strength of RFI sources are consistent, i.e., they are observed as coherent sources. This is in contrast with actual observations, which almost reach the thermal noise and are thought not to be limited by RFI. Therefore, we conclude that it is unlikely that there are undetected RFI sources that will become visible after long-time integration. Consequently, it is likely that RFI will not prevent an EoR detection with LOFAR.
\end{abstract}

>>>>

\begin{abstract}
Faint undetected sources of radio-frequency interference (RFI) might become visible after long time-integration of radio observations if they are consistently present over time. Thereby, they might obstruct the detection of the weak astronomical signals of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals from the time of the earliest structures in the Universe. We explore the RFI situation at 30--163~MHz by studying brightness histograms of visibility data observed with LOFAR, similar to radio-source-count analyses that are used in cosmology. The histograms show an RFI distribution that follows a power-law distribution with an estimated exponent around -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law implies that the current LOFAR EoR observations should be severely RFI limited if the integrated strength of RFI sources are consistent, i.e., they are observed as coherent sources. This is in contrast with actual observations, which almost reach the thermal noise and are thought not to be limited by RFI. Therefore, we conclude that it is unlikely that there are undetected RFI sources that will become visible after long-time integration. Consequently, it is likely that RFI will not prevent an EoR detection with LOFAR.
\end{abstract}
